\textbf{The Heart can Lie: The Role of Interoception and Theory of Mind in Deception}

### 2023-04-23 paper/manuscript.Rmd
add abstract

affiliation:
  - id            : "1"
    institution   : "School of Psychology, University of Sussex, UK"
  - id            : "2"
    institution   : "School of Social Sciences, Nanyang Technological University, Singapore"
  - id            : "3"
    institution   : "Centre for Research and Development in Learning, Nanyang Technological University, Singapore"
  - id            : "4"
    institution   : "Lee Kong Chian School of Medicine, Nanyang Technological University, Singapore"
  - id            : "5"
    institution   : "National Institute of Education, Nanyang Technological University, Singapore"
abstract: |
output:
  papaja::apa6_pdf:
    keep_tex: yes
keywords: "Deception, Interoception, Theory of Mind, Poygraph, Lying Ability"
wordcount: '3720'
bibliography: ["references.bib"]
floatsintext: yes
figsintext: yes
figurelist: no
tablelist: no
footnotelist: no
linenumbers: yes
mask: no
draft: no
documentclass: apa6
classoption: man
header-includes:
- \usepackage[labelfont=bf, font={color=gray,small}]{caption}
- \usepackage{float}
- \usepackage[document]{ragged2e}
csl: utils/apa.csl
---

@@

The analysis was not pre-registered (stemming out from a student's project), but the full reproducible analysis script, statistical results report, and data, are available at 
https://github.com/DominiqueMakowski/DeceptionInteroTom

@@

The material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, etc.) for this research is available at **https://github.com/DominiqueMakowski/DeceptionInteroTom**.

<!-- # Funding -->

<!-- **no funding** -->

@@

We would like to thank Jia Rong Low and Maximilian Yong for their help in data collection.

>>>>

affiliation:
  - id            : "1"
    institution   : "School of Psychology, University of Sussex, UK"
  - id            : "2"
    institution   : "School of Social Sciences, Nanyang Technological University, Singapore"
  - id            : "3"
    institution   : "Centre for Research and Development in Learning, Nanyang Technological University, Singapore"
  - id            : "4"
    institution   : "Lee Kong Chian School of Medicine, Nanyang Technological University, Singapore"
  - id            : "5"
    institution   : "National Institute of Education, Nanyang Technological University, Singapore"
abstract: |
  While a large part of the deception literature focuses on lying detection, the factors contributing to one's ability to lie remain unclear. The present study examined the contribution of Theory of Mind (ToM) and interoception on our ability to lie using a directed lie paradigm with two conditions ("Interrogation" and "Polygraph"), designed to enhance each of the two mechanisms. Using a mixed-design Bayesian analysis, we show that various facets of interoceptive abilities are positively related to the self-rated confidence in one's own lies, especially when under the belief that bodily signals are being monitored (i.e., in the "Polygraph" condition). Beyond providing evidence for the role of the body in lying and raising interesting questions for deception science, these results carry practical implications for criminology and lie detection protocols.
output:
  papaja::apa6_pdf:
    keep_tex: yes
keywords: "Deception, Interoception, Theory of Mind, Poygraph, Lying Ability"
wordcount: '3720'
bibliography: ["references.bib"]
floatsintext: yes
figsintext: yes
figurelist: no
tablelist: no
footnotelist: no
linenumbers: yes
mask: no
draft: no
documentclass: apa6
classoption: man
header-includes:
- \usepackage[labelfont=bf, font={color=gray,small}]{caption}
- \usepackage{float}
- \usepackage[document]{ragged2e}
csl: utils/apa.csl
---

@@

The analysis was not pre-registered (stemming out from a student's project), but the full reproducible analysis script, statistical results report, and data, are available at 
<!-- **[masked for blinding]** -->
**https://github.com/DominiqueMakowski/DeceptionInteroTom**

@@

The material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, etc.) for this research is available at 
**https://github.com/DominiqueMakowski/DeceptionInteroTom**.
<!-- **[masked for blinding]**. -->

@@

<!-- **[masked for blinding]** -->
We would like to thank Jia Rong Low and Maximilian Yong for their help in data collection.

### 2023-04-23 paper/manuscript.tex
add abstract

@@

The analysis was not pre-registered (stemming out from a student's project), but the full reproducible analysis script, statistical results report, and data, are available at
\url{https://github.com/DominiqueMakowski/DeceptionInteroTom}

>>>>

\abstract{%
While a large part of the deception literature focuses on lying detection, the factors contributing to one's ability to lie remain unclear. The present study examined the contribution of Theory of Mind (ToM) and interoception on our ability to lie using a directed lie paradigm with two conditions (``Interrogation'' and ``Polygraph''), designed to enhance each of the two mechanisms. Using a mixed-design Bayesian analysis, we show that various facets of interoceptive abilities are positively related to the self-rated confidence in one's own lies, especially when under the belief that bodily signals are being monitored (i.e., in the ``Polygraph'' condition). Beyond providing evidence for the role of the body in lying and raising interesting questions for deception science, these results carry practical implications for criminology and lie detection protocols.
}

@@

The analysis was not pre-registered (stemming out from a student's project), but the full reproducible analysis script, statistical results report, and data, are available at
\textbf{\url{https://github.com/DominiqueMakowski/DeceptionInteroTom}}
